---
title: "Steam Game Analytics R Package"
author: "Travis"
date: "2025-04-22"
output: html_document
---
====Overview====
For my final project, I would like to create a package that will allow a user
to view the data of the steam web page to allow visual analytics on the games.
I want functions that will allow the user to view specific games numbers and then
view the overall numbers on a histogram shiny app where the user may put in the 
variables and see where the overall list of games end up.

====HOW TO USE====
This part will be a lot of repeating code I have already stated below, but I will leave it here to make it easier for users to understand and use.

1. Use devtools::install() to downlaod package
2. load package using library(VideoGameAnalytics)
3. make sure to load all file by using devtools::load_all()
4. Set API key by using set_rawg_key("YOUR_API_KEY_HERE")
5. store data frame and how many rows you want using games_df <- fetch_game_data_env(pages = ###(amount of pages you want (1 page = 40)))
6. Use functions:
  *get_games()
  EX. games <- get_games(search  = "call of duty")
  
  *get_game_details()
  EX. outlast <- get_game_details(3790)
  
  *get_genres()
  
  *summarize_game_data(games_df)
  
7. Run shinyapp histogram: shiny::runApp("ShinyVG.R")
  
====Load Libraries====
```{r, load_libraries}
library(httr)
library(jsonlite)
library(dplyr)
library(shiny)
library(ggplot2)
```

====Load API====
```{r, load_api}
# Use this function to set API key once per session
set_rawg_key <- function(key) {
  Sys.setenv(RAWG_API_KEY = key)
}

# Main function to fetch game data (requires an API key)
fetch_game_data <- function(api_key = Sys.getenv("RAWG_API_KEY"), pages = 1, enrich = FALSE) {
  if (api_key == "") stop("API key not set. Use set_rawg_key().")
  all_games <- list()

  for (page in 1:pages) {
    url <- paste0(
      "https://api.rawg.io/api/games?key=", api_key,
      "&page=", page, "&page_size=40"
    )
    response <- GET(url)
    if (response$status_code != 200) stop("Failed to fetch data.")
    
    data <- fromJSON(content(response, "text"))
    games <- data$results

    if (enrich) {
      games <- lapply(games, function(game) {
        full <- tryCatch(get_game_details(game$id), error = function(e) NULL)
        if (!is.null(full)) game$genres <- full$genres
        game
      })
    }

    all_games[[page]] <- games
    Sys.sleep(1)
  }

  bind_rows(all_games) %>%
    select(name, released, rating, ratings_count, metacritic, playtime, genres) %>%
    filter(!is.na(rating))
}

# Friendly wrapper
fetch_game_data_env <- function(pages = 1, enrich = FALSE) {
  fetch_game_data(pages = pages, enrich = enrich)
}

```


====GAME FINDER====
Create a function where the user may search for a video game through a variety
of ways. EX.  games <- get_games(search  = "call of duty")
```{r}
get_games <- function(search = NULL, genres = NULL, platforms = NULL,
                      dates = NULL, ordering = NULL, page_size = 20) {
  parameters <- list(
    search = search,
    genres = genres,
    platforms = platforms,
    dates = dates,
    ordering = ordering,
    page_size = page_size
  )
  rawg_request("/games", parameters)$results
}

```

====FIND DETAILS====
Allow user to gather details of the game 

EX. outlast <- get_game_details(3790) 
    outlast$genre
    
```{r}
get_game_details <- function(game_id) {
  rawg_request(paste0("/games/", game_id))
}
```

====FIND GENRES====
Allow user to categorize the genres of the file
```{r}
get_genres <- function() {
  rawg_request("/genres")$results
}
```

====DATA SUMMARY====

Summarize data frame using:

> summarize_game_data(games_df)

```{r}
summarize_game_data <- function(games_df) {
  games_df %>%
    summarise(
      avg_rating = mean(rating, na.rm = TRUE),
      num_games = n(),
      earliest = min(released, na.rm = TRUE),
      latest = max(released, na.rm = TRUE)
    )
}
```

====SHINY APP====
Create a shiny app that will allow the user to create a histogram using the numerous variables
```{r}

#Tell user on what to download if the shiny app doesn't work
run_game_histogram_app <- function(data) {
  if (!requireNamespace("shiny", quietly = TRUE)) {
    stop("The 'shiny' package is required. Please install it with install.packages('shiny').")
  }
  if (!requireNamespace("ggplot2", quietly = TRUE)) {
    stop("The 'ggplot2' package is required. Please install it with install.packages('ggplot2').")
  }

#Create the shiny app (specifically histogram)
  shinyApp(
    ui = fluidPage(
      titlePanel("Game Metric Histogram"),
      sidebarLayout(
        sidebarPanel(
          selectInput("metric", "Choose metric to plot:",
                             choices = names(data)[sapply(data, is.numeric)],
                             selected = "rating"),
          sliderInput("bins", "Number of bins:", min = 5, max = 50, value = 20)
        ),
        mainPanel(
          plotOutput("histPlot")
        )
      )
    ),
    server = function(input, output) {
      output$histPlot <- renderPlot({
        ggplot(data, aes_string(x = input$metric)) +
          geom_histogram(bins = input$bins, fill = "steelblue", color = "black") +
          labs(
            title = paste("Histogram of", input$metric),
            x = input$metric, y = "Count"
          ) +
          theme_minimal()
      })
    }
  )
}
```

====RUN SHINY APP====
```{r}
games <- fetch_game_data(pages = 2, api_key = 'a402e3468b1640479e56d85cdf7e9e2c&page=')
run_game_histogram_app(games)
```


  
